---
title: "SeaBird PAR Kd calculations"
output: html_notebook
---

This R notebook is for calculating attentuation coefficients (Kd) in western Lake Erie for PAR wavelengths using SeaBird PAR profiles collected by Dack Stuart at CIGLR.  

Kd (units are inverse meters) is related to the decrease in irradiance in the water column with depth via the following equation: 
$$I_{\lambda,z} = I_{\lambda,0} * e^{-(Kd*z)} $$   

where $I_{\lambda,z}$ is the irradiance at a depth (z) for a given wavelength $(\lambda)$. $I_{\lambda,0}$ is the surface irradiance for a given.  

After rearranging the equation as follows, we can see that the slope of the natural log of irradiance at depth is equal to Kd:  

1. $\frac{I_{\lambda,z}}{I_{\lambda,0}} = e^{-(Kd*z)}$ (divide both sides by $I_{\lambda,0}$)  

2. $\ln(\frac{I_{\lambda,z}}{I_{\lambda,0}}) = -Kd*z$ (log transform both sides)   

3. $\ln({I_{\lambda,z}}) - \ln({I_{\lambda,0}}) = -Kd*z$ (rewrite the logarithmic expression)  

4. $\ln({I_{\lambda,z}}) = \ln({I_{\lambda,0}}) - Kd*z$  (add $\ln({I_{\lambda,0}})$ to both sides)  


The above equation has the form for the general equation of a line: $y = b - mx$  
So the slope of ln(irradiance) vs. depth at each date will give us Kd of each wavelength. 

Load the needed R packages:
```{r}
library(dplyr)
library(broom)
library(ggplot2)
library(patchwork)
```

First, import a data frame of spectral irradiance with depth in the PAR wavelengths:  
```{r}
SeaBird_df <- read.table("SeaBird_PAR_Transmission.txt", header=TRUE, sep="\t")
```

Log transform the irradiance data at each wavelength:
```{r}
#loop through 3 - number of columns in the dataframe
#Starting at 3 because column 1 is the date and column 2 is depth.
for (i in 3:dim(SeaBird_df)[2]){
  SeaBird_df[,i] <- log(SeaBird_df[,i])
}
```
Remove rows where NANs were produced:
```{r}
SeaBird_no_nans <- na.omit(SeaBird_df)
```

Find the slope of log transformed irradiance and depth for each wavelength (columns) on each date:  
```{r}
#Get a vector of dates to loop through:
dates <- unique(SeaBird_no_nans$Datetag)

#create an empty array to save the linear regressions in:  
#number of rows will be the product of length of dates and wavelength columns:  
kd_lm_results <- data.frame(matrix(ncol=3, nrow=length(3:length(SeaBird_no_nans)) * length(dates)))

#Loop through all the wavelengths in each date, regressing log transformed irradiance against depth:  
count <- 0 #Set up a counter, to keep track of the number of times we went through the loop and index the results later
for (date in dates){
  temp_df <- filter(SeaBird_no_nans, Datetag == date) #get only values from one date.
  #loop through each column of the dataframe that contains irradiance data. Start at column 3 because dates are in column 1 and depth is in column 2
  for (i in 3:length(SeaBird_no_nans)){ 
    count <- count + 1 
    kd_lm_results[count,1] <- date #store date
    kd_lm_results[count,2] <- colnames(SeaBird_no_nans)[i] #store wavelength
    kd_lm <- summary(lm(temp_df[ , i] ~ temp_df$Depth)) #regress ln(irradiance) vs depth
    kd_lm_results[count,3] <- kd_lm$coefficients[2,1] #store the slope value (Kd)
  }
}

#Rename column names in results table to something more useful:
colnames(kd_lm_results) <- c("Date", "Wavelength", "Kd")
```

Calculate average Kd values for each wavelength:  
```{r}
#First convert Kd data into positive numbers:
kd_lm_results$Kd <- abs(kd_lm_results$Kd)

Average_Kd <- kd_lm_results %>%
  group_by(Wavelength) %>%
  summarise(n=n(), Kd_avg=mean(Kd), Kd_SD=sd(Kd)) %>%
  mutate(Kd_95_CI=Kd_SD/sqrt(n)*1.96)

#Clean up wavelength entry for plotting:
Average_Kd$Wavelength <- gsub("X", "", Average_Kd$Wavelength)
``` 

Plot average Kd at each wavelength:
```{r}
ggplot(Average_Kd, aes(x=as.numeric(Wavelength), y=Kd_avg)) +
  geom_point() +
  geom_errorbar(aes(ymin=Kd_avg-Kd_95_CI, ymax=Kd_avg+Kd_95_CI)) +
  ggtitle("PAR attenuation") +
  theme_classic() +
  theme(plot.background = element_rect(color = "NA"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.x = element_line(size=0.2),
        axis.line.y = element_line(size=0.2),
        axis.ticks.length = unit(-0.1, "cm"),
        axis.text.x = element_text(angle = 45, hjust = 1, margin = margin(t = 5, r = 0, b = 0,)),
        axis.text.y = element_text(margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.y = element_text(margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.x = element_text(margin = margin(t = 5, r = 0, b = 0,))) +
  coord_cartesian(ylim=c(0.6,2.2)) +
  scale_x_continuous(breaks=seq(300,800, by=20)) +
  scale_y_continuous(breaks=seq(0.6,2.2, by=0.2)) +
  ylab(expression("Attenuation Coefficient (m"^-1*")")) +
  xlab("Wavelength (nm)")
```
Let's add in the UV attenuation data determined with the C-OPS:  
```{r}
#Import COPS dataframe:  
COPS_df <- read.table("COPS_df.txt", header=TRUE, sep="\t")

#Combine with the Seabird dataframe, keeping only values above 380 nm:
Average_Kd <- filter(Average_Kd, Wavelength > 390)
Combined_Seabird_COPS_df <- rbind(COPS_df, Average_Kd)
Combined_Seabird_COPS_df$Wavelength <- as.numeric(Combined_Seabird_COPS_df$Wavelength)
#Round the wavelength to the nearest whole number:
Combined_Seabird_COPS_df$Wavelength <- round(Combined_Seabird_COPS_df$Wavelength, digits = 0)

Kd_wave_plot <- ggplot(Combined_Seabird_COPS_df, aes(x=Wavelength, y=Kd_avg)) +
  geom_point(size=0.5) +
  geom_errorbar(aes(ymin=Kd_avg-Kd_95_CI, ymax=Kd_avg+Kd_95_CI), size=0.1) +
  theme_classic() +
  theme(plot.background = element_rect(color = "NA"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.x = element_line(size=0.2),
        axis.line.y = element_line(size=0.2),
        axis.ticks.length = unit(-0.05, "cm"),
        axis.text.x = element_text(size = 18, color = "black", angle = 45, hjust = 1,
                                   margin = margin(t = 5, r = 0, b = 0,)),
        axis.text.y = element_text(size = 18, color = "black", margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.y = element_text(size = 20, margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.x = element_text(size = 20, margin = margin(t = 5, r = 0, b = 0,))) +
  coord_cartesian(ylim=c(0.6,10)) +
  scale_x_continuous(breaks=seq(300,800, by=20)) +
  scale_y_continuous(breaks=seq(2,10, by=2)) +
  ylab(expression("Attenuation Coefficient (m"^-1*")")) +
  xlab("Wavelength (nm)")

Kd_wave_plot
```

From this curve, we will calculate the depth at which the fraction of light transmitted equals that transmitted through the bottles at each wavelength.   

First, let's plot the bottle + ND screen transmission:  
```{r}
#Import the dataframe:
Bottle_transmission_df <- read.table("Bottle_transmission.txt", header=TRUE, sep="\t")

Bottle_Transmission_plot <- ggplot(Bottle_transmission_df, aes(x=Wavelength, y=Film_Bottle_transmission)) +
  geom_line() +
  theme_classic() +
  theme(plot.background = element_rect(color = "NA"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.x = element_line(size=0.2),
        axis.line.y = element_line(size=0.2),
        axis.ticks.length = unit(-0.1, "cm"),
        axis.text.x = element_text(size = 18, color = "black", angle = 45, hjust = 1,
                                   margin = margin(t = 5, r = 0, b = 0,)),
        axis.text.y = element_text(size = 18, color = "black", margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.y = element_text(size = 20, margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.x = element_text(size = 20, margin = margin(t = 5, r = 0, b = 0,))) +
  scale_x_continuous(breaks=seq(200,600, by=100)) +
  ylab("Percent Transmission") +
  xlab("Wavelength (nm)")

Bottle_Transmission_plot
```

Merge the two dataframes, keeping only attenuation measurements that have a matching wavelength value in both the water column profiles and bottle transmission data:  
```{r}
Merged_df <- merge(Bottle_transmission_df, Combined_Seabird_COPS_df, by="Wavelength", all = FALSE)
```

Using the Kd values, calculate the depth at which $\frac{I_{\lambda,z}}{I_{\lambda,0}}$ is equal to bottle transmission during the experiments:  
```{r}
#Convert from percent back to a fraction:  
Merged_df$Film_Bottle_transmission <- Merged_df$Film_Bottle_transmission / 100
#Find the depth in the lake that correpsonds to the light exposure in the experiments:  
Merged_df$Cor_Lake_Depth <- log(Merged_df$Film_Bottle_transmission) / (Merged_df$Kd_avg * -1)
```

Plot the corresponding lake depth for each wavelength:  
```{r}
Cor_Lake_Depth_plot <- ggplot(Merged_df, aes(x=Wavelength, y=Cor_Lake_Depth)) +
  geom_line() +
  theme_classic() +
  theme(plot.background = element_rect(color = "NA"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line.x = element_line(size=0.2),
        axis.line.y = element_line(size=0.2),
        axis.ticks.length = unit(-0.05, "cm"),
        axis.text.x = element_text(size = 18, color = "black", angle = 45, hjust = 1,
                                   margin = margin(t = 5, r = 0, b = 0,)),
        axis.text.y = element_text(size = 18, color = "black", margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.y = element_text(size = 20, margin = margin(t = 0, r = 5, b = 0,)),
        axis.title.x = element_text(size = 20, margin = margin(t = 5, r = 0, b = 0,))) +
  coord_cartesian(ylim=c(0,4)) +
  scale_y_continuous(breaks=seq(0,4, by=0.5)) +
  scale_x_continuous(breaks=seq(300,800, by=20)) +
  ylab("Depth (m)") +
  xlab("Wavelength (nm)")

Cor_Lake_Depth_plot
```
Combine these three plots into one figure:  
```{r}
light_transmission_plot <- Kd_wave_plot + Bottle_Transmission_plot + Cor_Lake_Depth_plot + plot_layout(ncol=1)
light_transmission_plot
ggsave("light_transmission_plot.pdf", light_transmission_plot, width = 8, height = 18, units = "in", dpi=600)
```








